Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-05-18 scripts/MouseSexDiffApril20_2020.Rmd
Updates

---
title: "Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications"
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair
  Senior, Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  bookdown::html_document2: 
            toc: true
            code_folding: hide
            toc_depth: 6
            toc_float: yes
            number_sections: yes
            theme: flatly
  html_notebook:
    toc: yes
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

@@

Our supplementary material follows the general structure described in our methods and are outlined here in Figure \@ref(fig:FigG1). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis, and how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document or conviently referenecd quickly using the side table of contents panel. We will refer back to this figure as we describe the various aspects of the code and data processing.

@@

In this section, we cover the workflow described in Figure \@ref(fig:FigG1) C.

@@

@@

```{r loop, message=FALSE, warning=FALSE}

@@

Procedure names, grouping variables and trait names ("parameter_names") are merged back together with the results from the metafor analysis above.

@@

## Meta-analysis: condensing non-independent traits 
This section describes the workflow depicted in Figure \@ref(fig:FigG1) F.

### Dealing with Correlated Parameters, preparation

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted meta analyses on these correlated parameters to collapse the number of levels.

### Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

### Meta-analyses on correlated (sub-)traits, using `robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

## Second-order meta-analysis for functional groups
Here, we  describes the workflow depicted in Figure \@ref(fig:FigG1) H.

### Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR

#### Preparing the data

Nesting, calculating the number of parameters within each grouping term, and running the meta-analyses

@@

# **calculate number of parameters per grouping term

@@

  mutate(
    model_lnCVR = map(data, ~ metafor::rma.uni(
      yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    )),
    model_lnVR = map(data, ~ metafor::rma.uni(
      yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    )),
    model_lnRR = map(data, ~ metafor::rma.uni(
      yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    ))
  )

@@

  mutate(
    model_lnCVR = map(data, ~ metafor::rma.uni(
      yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    )),
    model_lnVR = map(data, ~ metafor::rma.uni(
      yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    )),
    model_lnRR = map(data, ~ metafor::rma.uni(
      yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
    ))
  )
```

#### Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

# Function will take the averall results and extract the relevant trait of interest
extract_trait <- function(data, trait){
  tmp <- data %>% 
        filter(., GroupingTerm == trait) %>% 
        mutate(lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se, lnCVR_I2 = .[[4]][[1]]$I2,
                lnVR = .[[5]][[1]]$b,  lnVR_lower = .[[5]][[1]]$ci.lb,  lnVR_upper = .[[5]][[1]]$ci.ub,  lnVR_se = .[[5]][[1]]$se,  lnVR_I2 = .[[5]][[1]]$I2,
                lnRR = .[[6]][[1]]$b,  lnRR_lower = .[[6]][[1]]$ci.lb,  lnRR_upper = .[[6]][[1]]$ci.ub,  lnRR_se = .[[6]][[1]]$se,  lnRR_I2 = .[[6]][[1]]$I2)  %>%
        select(., GroupingTerm, lnCVR:lnRR_I2)

@@

All <- overall_all1 %>% 
       mutate(lnCVR = .[[2]][[1]]$b, lnCVR_lower = .[[2]][[1]]$ci.lb, lnCVR_upper = .[[2]][[1]]$ci.ub, lnCVR_se = .[[2]][[1]]$se, lnCVR_I2 = .[[2]][[1]]$I2,
              lnVR = .[[3]][[1]]$b,  lnVR_lower = .[[3]][[1]]$ci.lb,  lnVR_upper = .[[3]][[1]]$ci.ub,  lnVR_se = .[[3]][[1]]$se,  lnVR_I2 = .[[3]][[1]]$I2,
              lnRR = .[[4]][[1]]$b,  lnRR_lower = .[[4]][[1]]$ci.lb,  lnRR_upper = .[[4]][[1]]$ci.ub,  lnRR_se = .[[4]][[1]]$se,  lnRR_I2 = .[[4]][[1]]$I2,)  %>%
       select(., lnCVR:lnRR_I2)

@@

# Visualising meta-analysis results
## Preparation for Figure 4 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

```{r hetero_loop, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, include=FALSE, cache = TRUE}

#The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). 
#As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

```{r hetero,eval=TRUE, echo = FALSE}
#Exclude traits without variation between mouse strains
#Merge data sets containing metafor results with procedure etc. names. Dealing with correlated parameters, and running of second -order meta-analysis for  heterogeneity data.

@@

### Preparation for Figure S2
#### Preparation for Figure S2 A
#This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

```{r FigS1, fig.height = 8, fig.width = 7, fig.cap=" Numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits (Phase 3, Figure 1.1)."}

@@

```{r Table4}
  kable(overall2, digits = 3, caption = "Summary of overall meta-analyses on the functional trait group level (GroupingTerm). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95% CI's, standard error and I2 values are provided.") %>%
   kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

>>>>

---
title: "Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications"
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair
  Senior, Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  bookdown::html_document2: 
            toc: true
            code_folding: hide
            toc_depth: 6
            toc_float: yes
            number_sections: yes
            theme: flatly
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

@@

This supplementary material follows the general structure described in our methods and is outlined here in Figure \@ref(fig:FigG1). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis. This iss followed by how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document. Various parts of the workflow can be conviently referenecd using the side table of contents panel. We will refer back to this figure as we describe the various aspects of the code and data processing in relevant places.

@@

In this section, we cover the workflow described in Figure \@ref(fig:FigG1) C and D.

@@

This section covers Figure \@ref(fig:FigG1) C, by conducting a meta-analsyis for all traits and producing the meta-analysed effect sizes (lnVR, lnCVR and lnRR).

@@

```{r loop, message=FALSE, warning=FALSE, result = "hide"}

@@

Here we cover Figure \@ref(fig:FigG1) D, removing non-coverging model results and non-informative traits.Procedure names, grouping variables and trait names ("parameter_names") are the merged back together with the results from the metafor analysis above.

@@

## Meta-analysis: condensing non-independent traits 
This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted meta analyses on these correlated parameters to collapse the number of levels. This section describes the workflow depicted in Figure \@ref(fig:FigG1) E & F.

### Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation. The workflow here relates to Figure \@ref(fig:FigG1) E.

@@

### Meta-analyses on correlated (sub-)traits, using `robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated. This section specifically undertakes the workflow described in Figure \@ref(fig:FigG1) F.

@@

## Second-order meta-analysis for functional groups
Here, we  describe the workflow depicted in Figure \@ref(fig:FigG1) H. We now use our uncorrelated traits from our first order meta-analysis to conduct a secondary meta-analysis by analysing the 9 functional trait groups. We esatimte an overall mean lnCVR, lnVR and lnRR for each functional trait group.

### Preparing the data

Nesting, calculating the number of parameters within each grouping term, and running the meta-analyses.

@@

# **Calculate number of parameters per grouping term

@@

  mutate(
    model_lnCVR = map(data, ~ metafor::rma.uni(
                      yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
                      control = list(optimizer = "optim", 
                                     optmethod = "Nelder-Mead", 
                                     maxit = 1000), 
                      verbose = F)),
    model_lnVR = map(data, ~ metafor::rma.uni(
                    yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                    control = list(optimizer = "optim", 
                                   optmethod = "Nelder-Mead", 
                                   maxit = 1000), 
                    verbose = F)),
    model_lnRR = map(data, ~ metafor::rma.uni(
                      yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                      control = list(optimizer = "optim", 
                                     optmethod = "Nelder-Mead", 
                                     maxit = 1000), 
                      verbose = F)))

@@

  mutate(
    model_lnCVR = map(data, ~ metafor::rma.uni(
                      yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
                      control = list(optimizer = "optim", 
                                     optmethod = "Nelder-Mead", 
                                     maxit = 1000), 
                      verbose = F)),
    model_lnVR = map(data, ~ metafor::rma.uni(
                    yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                    control = list(optimizer = "optim", 
                                   optmethod = "Nelder-Mead", 
                                   maxit = 1000), 
                    verbose = F)),
    model_lnRR = map(data, ~ metafor::rma.uni(
                    yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                    control = list(optimizer = "optim", 
                                   optmethod = "Nelder-Mead", 
                                   maxit = 1000), 
                    verbose = F)))
```

### Re-structuring the data for each grouping term
Here we delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

# Function will take the averall results and extract the relevant trait of interest
extract_trait <- function(data, trait){
  tmp <- data %>% 
        filter(., GroupingTerm == trait) %>% 
        mutate(      lnCVR = .[[4]][[1]]$b, 
               lnCVR_lower = .[[4]][[1]]$ci.lb, 
               lnCVR_upper = .[[4]][[1]]$ci.ub, 
                  lnCVR_se = .[[4]][[1]]$se, 
                  lnCVR_I2 = .[[4]][[1]]$I2,
                      lnVR = .[[5]][[1]]$b,  
                lnVR_lower = .[[5]][[1]]$ci.lb,  
                lnVR_upper = .[[5]][[1]]$ci.ub,  
                   lnVR_se = .[[5]][[1]]$se,  
                   lnVR_I2 = .[[5]][[1]]$I2,
                      lnRR = .[[6]][[1]]$b,  
                lnRR_lower = .[[6]][[1]]$ci.lb,  
                lnRR_upper = .[[6]][[1]]$ci.ub,  
                   lnRR_se = .[[6]][[1]]$se,  
                   lnRR_I2 = .[[6]][[1]]$I2)  %>%
        select(., GroupingTerm, lnCVR:lnRR_I2)

@@

All <- overall_all1 %>% 
       mutate(      lnCVR = .[[2]][[1]]$b, 
              lnCVR_lower = .[[2]][[1]]$ci.lb, 
              lnCVR_upper = .[[2]][[1]]$ci.ub, 
                 lnCVR_se = .[[2]][[1]]$se, 
                 lnCVR_I2 = .[[2]][[1]]$I2,
                     lnVR = .[[3]][[1]]$b,  
               lnVR_lower = .[[3]][[1]]$ci.lb,  
               lnVR_upper = .[[3]][[1]]$ci.ub,  
                  lnVR_se = .[[3]][[1]]$se,  
                  lnVR_I2 = .[[3]][[1]]$I2,
                     lnRR = .[[4]][[1]]$b,  
               lnRR_lower = .[[4]][[1]]$ci.lb,  
               lnRR_upper = .[[4]][[1]]$ci.ub,  
                  lnRR_se = .[[4]][[1]]$se,  
                  lnRR_I2 = .[[4]][[1]]$I2,)  %>%
       select(., lnCVR:lnRR_I2)

@@

## Second-order meta-analysis of heterogenity

The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses and specifically Figure \@ref(fig:FigG1) I & J. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

```{r hetero_loop, echo = TRUE, eval = TRUE, message=FALSE, warning=FALSE, cache = TRUE}

Here we exclude traits without variation between mouse strains and merge data sets containing metafor results with procedure etc. names. Dealing with the correlated parameters, and running the second -order meta-analysis for heterogeneity data (Figure \@ref(fig:FigG1) I)

```{r hetero,eval=TRUE, echo = TRUE}

# Meta-analysis results
## Preparation for Figure 4 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

@@

#This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

```{r FigS1, fig.height = 8, fig.width = 7, fig.cap=" Numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits."}

@@

```{r Table4}
  kable(overall2, digits = 3, caption = "Summary of overall meta-analyses on the functional trait group level (GroupingTerm). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95 percent CI's, standard error and I2 values are provided.") %>%
   kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```
